Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-06-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Start adding more research questions for us to answer
I'm adding these in Results as I think of them. We can add more, think about whether the ones we have are useful, and think about how to answer ones that we think are useful
Add some results

```{r echo = FALSE}
library(hurricaneexposure)
library(knitr)
library(ggplot2)
library(gridExtra)
```

@@

@@

Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000: direct hurricane-related deaths). Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000]. Some studies suggest methods to determine exposure, including inland exposure, to tropical storms at the county-level over many years [@Zandbergen2009]. Here, we offer tools, through the open-source `hurricaneexposure` package, to determine county-level exposure to tropical storms using several different metrics (distance, rainfall, wind). In this paper, we describe how this software package assesses exposure to tropical storms at the county level, provide an analysis of the advantages and disadvantages of different strategies for assessing county-level exposure, and explore variation in exposure assessments based on using different exposure metrics.

@@

Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies: @Grabich2016; @Zandbergen2009]. Here, we offer methods to map and output historic exposure to tropical storms that does not require the use of proprietary software but instead uses a package written in the R statistical programming language, which is free and open-source.

@@

#### Are exposure patterns different when using distance-, rain-, and wind-based metrics?

@@

@@

However, these best tracks should be fairly reliable for more recent years of storms, as we use here. Many of the uncertainties related to storm positions in best track data are more of a concern for the years before ... (e.g., pre-19[xx]).

@@

Rain from tropical storms can bring important risks to human health and property, even in counties well inland from the coast. For example, most of the direct hurricane-related deaths in the US from 1970 to 1999 were caused by freshwater flooding and were in inland, rather than coastal, counties [@Rappaport2000]. Rains from tropical storms can flood roads [@Rappaport2000].

@@

Several storm characteristics are associated with the likelihood of heavy rain and flooding from the storm, including the storm's forward speed and size [@Rappaport2000]. There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season, when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000].

@@

A tropical storm's high winds can bring a number of dangers, including health risks and property damage caused by structural damage of houses and other buildings, falling trees, and wind-borne debris [@Rappaport2000]. While severe winds pose an important threat, risks of a tropical storm can exist without severe wind; for example, one study found that most of the direct hurricane-related deaths in the US between 1970 and 1999 occurred in cases when wind was below hurricane strength (i.e., < 64 knots) [@Rappaport2000].

>>>>

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(hurricaneexposure)
library(knitr)
library(ggplot2)
library(gridExtra)
library(dplyr)
```

@@

Tropical storms bring a number of threats, including high winds, heavy rain and flooding, storm surge, tornadoes, ... . To determine whether a county is exposed to a particular tropical storm, a researcher could use criteria based on measurements of any of these values during the storm, use a metric based on the distance from the county to the storm's track, or use a metric that combines several of these measurements.

While the risks to life from some tropical storm threats (e.g., storm surge) have been greatly reduced over the past century, other risks (e.g., inland flooding) have remained more steady [@Rappaport2000]. Determining the metrics of storm exposure that are most associated with loss of life and property damage may help identify and emphasize the important threats that remain from tropical storms in the United States, which could allow similar future success in reducing these threats through community planning, warning systems, etc.

@@

Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000: direct hurricane-related deaths). Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000]. Some studies suggest methods to determine exposure, including inland exposure, to tropical storms at the county-level over many years [@Zandbergen2009].

Here, we offer tools, through the open-source `hurricaneexposure` package, to determine county-level exposure to tropical storms using several different metrics (distance, rainfall, wind). In this paper, we describe how this software package assesses exposure to tropical storms at the county level, provide an analysis of the advantages and disadvantages of different strategies for assessing county-level exposure, and explore variation in exposure assessments based on using different exposure metrics.

@@

Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies: @Grabich2016; @Zandbergen2009; @Czajkowski2011]. Here, we offer methods to map and output historic exposure to tropical storms that does not require the use of proprietary software but instead uses a package written in the R statistical programming language, which is free and open-source.

@@

#### Are exposure patterns different when using distance-, rain-, and wind-based metrics? In particular, (1) do some exposure metrics result in more "exposed" classifications than others and (2) are geographic patterns of exposure different among the different metrics?

Map of county exposures to the top 1000 county-storm pairs by distance-based exposure:

```{r echo = FALSE}
# Code for this map here
```

Map of county exposures to the top 1000 county-storm pairs by wind-based exposure:

```{r echo = FALSE}
# Code for this map here
```

Map of county exposures to the top 1000 county-storm pairs by rain-based exposure:

```{r echo = FALSE}
# Code for this map here
```

Map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this? What seems reasonable?] kilometers or less:

```{r echo = FALSE}
# Code for this map here
```

Map of total count of "exposures" by county, based on "exposed" = rainfall over a three-day window of 75 millimeters or more and distance of 500 kilometers or less:

```{r echo = FALSE}
# Code for this map here
```

@@

Here is a table of the top 10 storms by number of counties exposed, based on a distance metric of within 100 kilometers or less of the storm track:

```{r echo = FALSE, message = FALSE, warning = FALSE}
data("closest_dist")
all_counties <- unique(closest_dist$fips)

all_distance <- county_distance(counties = all_counties, start_year = 1988, 
                        end_year = 2011, dist_limit = 60)

distance_exposed <- all_distance %>%
        group_by(storm_id) %>%
        summarize(number_of_counties_exposed = n()) %>%
        arrange(desc(number_of_counties_exposed)) %>%
        top_n(10)

distance_exposed %>%
        mutate(storm_id = gsub("-", " (", storm_id),
               storm_id = paste0(storm_id, ")")) %>%
        kable(col.names = c("Storm", "# of exposed counties"),
              caption = "Top 10 storms based on the number of counties exposed, using a metric of distance.")
```

Here are plots of the top four storms based on number of counties exposed by distance:

```{r echo = FALSE, fig.width = 10, fig.height = 8}
top_distance_exposed <- function(storm_id){
        my_map <- map_distance_exposure(storm_id, dist_limit = 60)
        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
                             plot_points = FALSE) + 
                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
}

n <- 4
distance_plots <- lapply(distance_exposed$storm_id[1:n], top_distance_exposed)
names(distance_plots) <- letters[1:n]
grid.arrange(distance_plots$a, distance_plots$b, distance_plots$c,
             distance_plots$d, ncol = 2)
```

Here is a table of the top 10 storms by number of counties exposed, based on a rain metric of 75 millimeters or more and within 500 kilometers or less of the storm track:

```{r echo = FALSE, message = FALSE, warning = FALSE}
data("closest_dist")
all_counties <- unique(closest_dist$fips)

all_rain <- county_rain(counties = all_counties, start_year = 1988, 
                        end_year = 2011, rain_limit = 75, dist_limit = 500,
                        days_included = -1:1)

rain_exposed <- all_rain %>%
        group_by(storm_id) %>%
        summarize(number_of_counties_exposed = n()) %>%
        arrange(desc(number_of_counties_exposed)) %>%
        top_n(10)

rain_exposed %>%
        mutate(storm_id = gsub("-", " (", storm_id),
               storm_id = paste0(storm_id, ")")) %>%
        kable(col.names = c("Storm", "# of exposed counties"),
              caption = "Top 10 storms based on the number of counties exposed, using a metric of rainfall.")
```

Here are plots of the exposed counties for the top 4 storms based on that rain metric:

```{r echo = FALSE, fig.width = 10, fig.height = 8}
top_rain_exposed <- function(storm_id){
        my_map <- map_rain_exposure(storm_id, rain_limit = 75, 
                                    dist_limit = 500)
        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
                             plot_points = FALSE) + 
                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
}

n <- 4
rain_plots <- lapply(rain_exposed$storm_id[1:n], top_rain_exposed)
names(rain_plots) <- letters[1:n]
grid.arrange(rain_plots$a, rain_plots$b, rain_plots$c, rain_plots$d, ncol = 2)
```

@@

However, these best tracks should be fairly reliable for more recent years of storms, as we use here. Many of the uncertainties related to storm positions in best track data are more of a concern for the years before ... (e.g., pre-19[xx], before aircraft reconnaisance of Atlantic basin tropical storms).

@@

Rain from tropical storms can bring important risks to human health and property, even in counties well inland from the coast. For example, most of the direct hurricane-related deaths in the US from 1970 to 1999 were caused by freshwater flooding and were in inland, rather than coastal, counties [@Rappaport2000]. Rains from tropical storms can flood roads [@Rappaport2000]. Examples of historical tropical storms that caused many direct deaths from flooding in inland counties include Hurricane Diane in 1955 and Hurricane Camille in 1969 [@Rappaport2000]; more recent examples include Hurricane Floyd in 1999, Hurricane Alberto in 1994, Hurricane Charley in 1998, and Hurricane Fran in 1996 [@Rappaport2000].

@@

Several storm characteristics are associated with the likelihood of heavy rain and flooding from the storm, including the storm's forward speed and size [@Rappaport2000]. There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season (e.g., June--August), when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000].

@@

A tropical storm's high winds can bring a number of dangers, including health risks and property damage caused by structural damage of houses and other buildings, falling trees, and wind-borne debris [@Rappaport2000]. These types of wind-related damage can occur when winds are below hurricane strength (i.e., < 64 knots) [@Rappaport2000]. While severe winds pose an important threat, risks of a tropical storm can exist without severe wind; for example, one study found that most of the direct hurricane-related deaths in the US between 1970 and 1999 occurred in cases when wind was below hurricane strength, including for Tropical Storms Charley in 1998 and Alberto in 1994 [@Rappaport2000]. Further, wind-based exposure metrics are directly linked to a storm's strength, and some research suggests that hurricane strength and landfall is not strongly associated with the number of direct deaths ultimately caused by the storm [@Rappaport2000].

### 2016-06-18 writing/DraftExposurePaper/DraftExposurePaper.tex
Add some results

@@

Many outcomes of interest are available at county-level aggregations
(e.g., counts of health outcomes: Grabich et al. (2016): premature
births and low birth weights; Grabich et al. (2015): birth rates).
Further, decisions and policies to prepare for and respond to storms are
often undertaken at the county level (Zandbergen 2009). Some studies
suggest methods to determine exposure, including inland exposure, to
tropical storms at the county-level over many years (Zandbergen 2009).
Here, we offer tools, through the open-source \texttt{hurricaneexposure}
package, to determine county-level exposure to tropical storms using
several different metrics (distance, rainfall, wind). In this paper, we
describe how this software package assesses exposure to tropical storms
at the county level, provide an analysis of the advantages and
disadvantages of different strategies for assessing county-level
exposure, and explore variation in exposure assessments based on using
different exposure metrics.

Current hurricane exposure assignments are often nonspecific {[}GBA-- I
found this wording a bit unclear{]} and may lead to missclassification
of exposure at the county-level (Grabich et al. 2016). Researchers have
found heterogeneity in exposure assignments of counties, both within and
between storms, when different metrics of exposure are used (Grabich et
al. 2016; Grabich et al. 2015). Such missclassification can have
important implications of assessments of the public health and economic
impacts of tropical storms.

Many of the studies that have assessed exposure to tropical storms in
the US have used geographical information system software (e.g., ArcGIS)
(example studies: Grabich et al. 2016; Zandbergen 2009). Here, we offer
methods to map and output historic exposure to tropical storms that does
not require the use of proprietary software but instead uses a package
written in the R statistical programming language, which is free and
open-source.

@@

If exposure to tropical storms over multiple storms and years can be
assessed, these exposure datasets can be joined with other time series
to explore the impacts of tropical storms. For example, daily counts of
human health outcomes in environmental epidemiology studies are often
available aggregated at the county level, and such data has often been
paired with time series of environmental exposures (e.g., air pollution,
temperature) to determine associated risks.

@@

We collected ``best tracks'' data on hurricane tracks for Atlantic basin
storms between 1988 and 2014 from the extended best tracks database.
This dataset is based on a poststorm assessment of each storm conducted
by the United States National Hurricane Center (NHC) and incorporates
data from a variety of sources, including satellite data and, when
available, aircraft reconnaisance data (Landsea and Franklin 2013). This
data gives time stamps for each observation in Coordinated Universal
Time (UTC; also known as Zulu Time, sometimes indicated by ``Z'').

@@

We calculated the distance between each county's center and each of the
15-minute-interval estimates of a storm's location. To do this, we used
the Great Circle method to calculate the distance between pairs of
latitude and longitude coordinates, using the R package \texttt{sp}. For
each county, we used the population mean center, based on the U.S.
Census's 2010 Decennial Census, as the center location of the county.
From these distance calculations, we identified the date-time and
distance between the storm track and the county center at the 15-minute
interval when the storm was closest to the county center. We used the
\texttt{countytimezones} R package to convert each date-time of the
closest storm approach for a county to the county's local time zone and
identified the local date when the storm was closest to the county. This
``closest date'' was used to pair the distance estimates with rainfall
estimates.

@@

NLDAS does not provide information on precipitation over the ocean,
preventing its use in assessing the develpment of the storm prior to
landfall (Villarini et al. 2011). This point is moot for this analysis.

@@

@@

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-2-1.pdf}

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

However, these best tracks should be fairly reliable for more recent
years of storms, as we use here. Many of the uncertainties related to
storm positions in best track data are more of a concern for the years
before \ldots{} (e.g., pre-19{[}xx{]}).

@@

There are a number of limitations to using distance to assess exposure
to tropical storms. Tropical storms vary in size and intensity, and a
measure of distance from the storm track will not incorporate these
differences and so could misclassify exposure both in terms of
generating false positives (counties close to the storm track of very
mild or small-radius storms) and false negatives (e.g., during very
large or intense storms). In particular, the forces of a storm tend to
be distributed around the center in a non-symmetrical way
{[}citation{]}, while distance-based exposure metrics that use buffers
tend to use an equal buffer distance on each side of the storm track
(e.g., {[}citations{]}).

@@

@@

Exposure classification based on rainfall has some advantages. For
example, it allows the identification of exposed counties that are
inland, rather than coastal, but that were affected by heavy rainfall
during the storm. Often, storm-related deaths are associated with inland
flooding. In fact, over half of the hurricane-related fatalities from
1970 to 1999 were a result of freshwater flooding, accounting for the
vast majority of deaths in inland counties (Rappaport 2000). In another
analysis, researchers found that 79\% of freshwater-drowning fatalities
occured in inland counties,(Czajkowski, Simmons, and Sutter 2011)
further stressing the importance of rainfall in hurricane risk analysis.
Storms can produce a lot of rain especially in certain topographies,
like near mountains, so counties that are well inland can sometimes
experience more extreme rain that other counties at similar distances
from the storm's track.

@@

@@

>>>>

Tropical storms bring a number of threats, including high winds, heavy
rain and flooding, storm surge, tornadoes, \ldots{} . To determine
whether a county is exposed to a particular tropical storm, a researcher
could use criteria based on measurements of any of these values during
the storm, use a metric based on the distance from the county to the
storm's track, or use a metric that combines several of these
measurements.

While the risks to life from some tropical storm threats (e.g., storm
surge) have been greatly reduced over the past century, other risks
(e.g., inland flooding) have remained more steady (Rappaport 2000).
Determining the metrics of storm exposure that are most associated with
loss of life and property damage may help identify and emphasize the
important threats that remain from tropical storms in the United States,
which could allow similar future success in reducing these threats
through community planning, warning systems, etc.

@@

Many outcomes of interest are available at county-level aggregations
(e.g., counts of health outcomes: Grabich et al. (2016): premature
births and low birth weights; Grabich et al. (2015): birth rates;
Rappaport (2000): direct hurricane-related deaths). Further, decisions
and policies to prepare for and respond to storms are often undertaken
at the county level (Zandbergen 2009; Rappaport 2000). Some studies
suggest methods to determine exposure, including inland exposure, to
tropical storms at the county-level over many years (Zandbergen 2009).

Here, we offer tools, through the open-source \texttt{hurricaneexposure}
package, to determine county-level exposure to tropical storms using
several different metrics (distance, rainfall, wind). In this paper, we
describe how this software package assesses exposure to tropical storms
at the county level, provide an analysis of the advantages and
disadvantages of different strategies for assessing county-level
exposure, and explore variation in exposure assessments based on using
different exposure metrics.

Current hurricane exposure assignments are often nonspecific {[}GBA-- I
found this wording a bit unclear{]} and may lead to missclassification
of exposure at the county-level (Grabich et al. 2016). Researchers have
found heterogeneity in exposure assignments of counties, both within and
between storms, when different metrics of exposure are used (Grabich et
al. 2016; Grabich et al. 2015). Such missclassification can have
important consequences in assessments of the public health and economic
impacts of tropical storms.

Many of the studies that have assessed exposure to tropical storms in
the US have used geographical information system software (e.g., ArcGIS)
(example studies: Grabich et al. 2016; Zandbergen 2009; Czajkowski,
Simmons, and Sutter 2011). Here, we offer methods to map and output
historic exposure to tropical storms that does not require the use of
proprietary software but instead uses a package written in the R
statistical programming language, which is free and open-source.

@@

If exposure to tropical storms over multiple storms and years can be
assessed, these exposure datasets can be joined with other time series
to explore the impacts of tropical storms. For example, daily counts of
human health outcomes in environmental epidemiology studies are often
available aggregated at the county level, and such data has often been
paired with time series of environmental exposures (e.g., air pollution,
temperature) to determine associated risks. Databases of storm impacts,
including direct deaths (Rappaport 2000), \ldots{}, and \ldots{}, are
often aggregated at the county level.

@@

We collected ``best tracks'' data on hurricane tracks for Atlantic basin
storms between 1988 and 2014 from the extended best tracks database.
This dataset is based on a post-storm assessment of each storm conducted
by the United States National Hurricane Center (NHC) and incorporates
data from a variety of sources, including satellite data and, when
available, aircraft reconnaisance data (Landsea and Franklin 2013). This
data gives time stamps for each observation in Coordinated Universal
Time (UTC; also known as Zulu Time, sometimes indicated by ``Z'').

@@

We calculated the distance between each county's center and each of the
15-minute-interval estimates of a storm's location. To do this, we used
the Great Circle method to calculate the distance between pairs of
latitude and longitude coordinates, using the R package \texttt{sp}. For
each county, we used the population mean center, based on the U.S.
Census's 2010 Decennial Census, as the center location of the county.
From these distance calculations, we identified the date-time and
distance between the storm track and the county center at the 15-minute
interval when the storm was closest to the county center. We identified
the date when the storm was closest to each county and converted this
date-time to the county's local time zone using the
\texttt{countytimezones} R package {[}citation{]}. This ``closest date''
was used to pair the distance estimates with rainfall estimates and can
also be used to merge exposure datasets with daily aggregated outcomes
like daily county-level mortality and morbidity counts.

@@

NLDAS does not provide information on precipitation over the ocean,
preventing its use in assessing the development of the storm prior to
landfall (Villarini et al. 2011). This point is moot for this analysis.

@@

\paragraph{\texorpdfstring{Are exposure patterns different when using
distance-, rain-, and wind-based metrics? In particular, (1) do some
exposure metrics result in more ``exposed'' classifications than others
and (2) are geographic patterns of exposure different among the
different
metrics?}{Are exposure patterns different when using distance-, rain-, and wind-based metrics? In particular, (1) do some exposure metrics result in more exposed classifications than others and (2) are geographic patterns of exposure different among the different metrics?}}\label{are-exposure-patterns-different-when-using-distance--rain--and-wind-based-metrics-in-particular-1-do-some-exposure-metrics-result-in-more-exposed-classifications-than-others-and-2-are-geographic-patterns-of-exposure-different-among-the-different-metrics}

Map of county exposures to the top 1000 county-storm pairs by
distance-based exposure:

Map of county exposures to the top 1000 county-storm pairs by wind-based
exposure:

Map of county exposures to the top 1000 county-storm pairs by rain-based
exposure:

Map of total count of ``exposures'' by county, based on ``exposed'' =
distance of 100 {[}or do we want a different distance limit for this?
What seems reasonable?{]} kilometers or less:

Map of total count of ``exposures'' by county, based on ``exposed'' =
rainfall over a three-day window of 75 millimeters or more and distance
of 500 kilometers or less:

@@

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-7-1.pdf}

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-8-1.pdf}

\paragraph{Which storms exposed the most counties when using distance-,
rain-, and wind-based metrics? What were the top county-storm exposures
for rain- and wind-based
metrics?}\label{which-storms-exposed-the-most-counties-when-using-distance--rain--and-wind-based-metrics-what-were-the-top-county-storm-exposures-for-rain--and-wind-based-metrics}

Here is a table of the top 10 storms by number of counties exposed,
based on a distance metric of within 100 kilometers or less of the storm
track:

\begin{longtable}[]{@{}lr@{}}
\caption{Top 10 storms based on the number of counties exposed, using a
metric of distance.}\tabularnewline
\toprule
Storm & \# of exposed counties\tabularnewline
\midrule
\endfirsthead
\toprule
Storm & \# of exposed counties\tabularnewline
\midrule
\endhead
Beryl (1994) & 195\tabularnewline
Frances (2004) & 194\tabularnewline
Cindy (2005) & 189\tabularnewline
Dennis (2005) & 187\tabularnewline
Chris (1988) & 181\tabularnewline
Erin (1995) & 175\tabularnewline
Allison (2001) & 170\tabularnewline
Isidore (2002) & 167\tabularnewline
Jeanne (2004) & 167\tabularnewline
Ivan (2004) & 159\tabularnewline
\bottomrule
\end{longtable}

Here are plots of the top four storms based on number of counties
exposed by distance:

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-10-1.pdf}

Here is a table of the top 10 storms by number of counties exposed,
based on a rain metric of 75 millimeters or more and within 500
kilometers or less of the storm track:

\begin{longtable}[]{@{}lr@{}}
\caption{Top 10 storms based on the number of counties exposed, using a
metric of rainfall.}\tabularnewline
\toprule
Storm & \# of exposed counties\tabularnewline
\midrule
\endfirsthead
\toprule
Storm & \# of exposed counties\tabularnewline
\midrule
\endhead
Frances (2004) & 454\tabularnewline
Ivan (2004) & 405\tabularnewline
Isidore (2002) & 307\tabularnewline
Floyd (1999) & 294\tabularnewline
Lee (2011) & 279\tabularnewline
Opal (1995) & 266\tabularnewline
Jeanne (2004) & 255\tabularnewline
Gustav (2008) & 252\tabularnewline
Ike (2008) & 236\tabularnewline
Katrina (2005) & 232\tabularnewline
\bottomrule
\end{longtable}

Here are plots of the exposed counties for the top 4 storms based on
that rain metric:

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-12-1.pdf}

\paragraph{Are storms with heavy rain exposure likely to occur early in
the hurricane
season?}\label{are-storms-with-heavy-rain-exposure-likely-to-occur-early-in-the-hurricane-season}

@@

However, these best tracks should be fairly reliable for more recent
years of storms, as we use here. Many of the uncertainties related to
storm positions in best track data are more of a concern for the years
before \ldots{} (e.g., pre-19{[}xx{]}, before aircraft reconnaisance of
Atlantic basin tropical storms).

@@

There are a number of limitations to using distance to assess exposure
to tropical storms. Tropical storms vary in size and intensity, and a
measure of distance from the storm track will not incorporate these
differences and so could misclassify exposure both in terms of
generating false positives (counties close to the storm track of very
mild or small-radius storms) and false negatives (e.g., during very
large or intense storms). While a number of storm characteristics are
strongly associated with distance from the storm's center (e.g., wind
and, at the coast, storm surge and waves; Rappaport (2000)), other
characteristics like rain can occur at dangerous levels well away from
the storm's central track (Rappaport 2000). Although storm surge was a
major cause of hurricane-related deaths in the past, more recently
freshwater flooding has become a much more common cause of deaths
directly resulting from tropical storms in the US (Rappaport 2000).

Further, the forces of a storm tend to be distributed around the center
in a non-symmetrical way {[}citation{]}, while distance-based exposure
metrics that use buffers tend to use an equal buffer distance on each
side of the storm track (e.g., {[}citations{]}).

@@

Rain from tropical storms can bring important risks to human health and
property, even in counties well inland from the coast. For example, most
of the direct hurricane-related deaths in the US from 1970 to 1999 were
caused by freshwater flooding and were in inland, rather than coastal,
counties (Rappaport 2000). Rains from tropical storms can flood roads
(Rappaport 2000). Examples of historical tropical storms that caused
many direct deaths from flooding in inland counties include Hurricane
Diane in 1955 and Hurricane Camille in 1969 (Rappaport 2000); more
recent examples include Hurricane Floyd in 1999, Hurricane Alberto in
1994, Hurricane Charley in 1998, and Hurricane Fran in 1996 (Rappaport
2000).

@@

Exposure classification based on rainfall has some advantages. For
example, it allows the identification of exposed counties that are
inland, rather than coastal, but that were affected by heavy rainfall
during the storm. Often, storm-related deaths are associated with inland
flooding. In the United States, over half of hurricane-related direct
deaths from 1970 to 1999 from Atlantic basin storms were a result of
freshwater flooding; as a result, most of these deaths were in inland,
rather than coastal, counties (Rappaport 2000). In another analysis,
researchers found that 79\% of freshwater-drowning fatalities occured in
inland counties,(Czajkowski, Simmons, and Sutter 2011) further stressing
the importance of rainfall in hurricane risk analysis. Storms can
produce a lot of rain especially in certain topographies, like near
mountains, so counties that are well inland can sometimes experience
more extreme rain that other counties at similar distances from the
storm's track.

@@

Several storm characteristics are associated with the likelihood of
heavy rain and flooding from the storm, including the storm's forward
speed and size (Rappaport 2000). There is some evidence that the worst
tropical storms, in terms of heavy rainfall and freshwater flood-related
deaths in the US, tend to occur early in the hurricane season (e.g.,
June--August), when it is more common to get the meteorological
conditions (weak steering currents) that lead to slow-moving storms
(Rappaport 2000).

@@

A tropical storm's high winds can bring a number of dangers, including
health risks and property damage caused by structural damage of houses
and other buildings, falling trees, and wind-borne debris (Rappaport
2000). These types of wind-related damage can occur when winds are below
hurricane strength (i.e., \textless{} 64 knots) (Rappaport 2000). While
severe winds pose an important threat, risks of a tropical storm can
exist without severe wind; for example, one study found that most of the
direct hurricane-related deaths in the US between 1970 and 1999 occurred
in cases when wind was below hurricane strength, including for Tropical
Storms Charley in 1998 and Alberto in 1994 (Rappaport 2000). Further,
wind-based exposure metrics are directly linked to a storm's strength,
and some research suggests that hurricane strength and landfall is not
strongly associated with the number of direct deaths ultimately caused
by the storm (Rappaport 2000).
